# Multiple Regression



```{r include=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(ggplot2)
library(factoextra)
library(broom)
```


```{r include=FALSE}
#Weather data - July 2018 to February 2019 (daily summaries)
changi_daily_weather <- read_excel(here::here("Data/daily_weather_changi.xlsx")) %>% 
  mutate(date=date(date))

# Changi traffic data - July 2018 to February 2019 (hourly data)
changi_stats <- read_excel(here::here("Data/changi_stats.xlsx"))

#Changi Weather and Traffic data - July 2018 to February 2019 (hourly data)
changi_hourly_weather<-read_excel(here::here("Data/weather_june18_feb19.xlsx")) 


changi_hourly_weather <-changi_hourly_weather %>% 
  mutate(total_tfc = departure_tfc+arrival_tfc)

```



```{r include=FALSE}
changi_stats <- read_excel(here::here("Data/changi_stats.xlsx")) %>% 
  mutate(daily_tfc = departure_tfc+arrival_tfc) %>% 
  mutate(date = date(local_time))


daily_tfc_flow<-changi_stats %>% 
  group_by(date) %>% 
  summarise(daily_tfc = sum(daily_tfc))

daily_stats_weather <- left_join(changi_daily_weather, daily_tfc_flow)
```


```{r include = FALSE}

hourly_stats_weather <-changi_hourly_weather %>% 
  mutate(hour = hour(local_time)) %>% 
  mutate(date = date(local_time)) %>% 
  mutate(temp = temp/10) %>% 
  mutate(temp_min = temp_min/10) %>% 
  mutate(temp_max = temp_max/10) %>% 
  select(-local_time)

```


```{r include = FALSE}
daily_stats_weather<- daily_stats_weather %>%
  mutate(daily_wind_speed = wind_speed) %>% 
  mutate(daily_max_wind_speed = max_wind_speed) %>%
  select(-wind_speed, - max_wind_speed)
  
data <-left_join(hourly_stats_weather, daily_stats_weather, by = "date")
```


We have two different datasets - a summary of the entire day's traffic and weather conditions as well as the hourly version which similarly looks at each hour's traffic and weather variables. The hourly data provides us with more details about the relationship between Changi weather and air traffic flows. 

In the following multiple regression models, we use wind speed, rainfall and temperature to determine which variables have the most impact on overall daily air traffic in Changi. 
```{r}
ols_daily_traffic_weather <- lm(daily_tfc ~ daily_rainfall_mm + daily_wind_speed + temp_mean_daily + `30min_rainfall_mm`  + `60min_rainfall_mm` + `120min_rainfall_mm`, daily_stats_weather)

glance(ols_daily_traffic_weather)

tidy(ols_daily_traffic_weather)
```



```{r}
ols_hourly_traffic_weather <- lm(total_tfc ~rainfall_mm + temp + humidity + wind_speed +pressure, hourly_stats_weather)

glance(ols_hourly_traffic_weather)

tidy(ols_hourly_traffic_weather)
```

##Analysis
