# Multidimensional Scaling

```{r include=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(ggplot2)
library(factoextra)
library(broom)
```

```{r include=FALSE}
#Weather data - July 2018 to February 2019 (daily summaries)
changi_daily_weather <- read_excel(here::here("Data/daily_weather_changi.xlsx")) %>% 
  mutate(date=date(date))

# Changi traffic data - Julu 2018 to February 2019 (hourly data)
changi_stats <- read_excel(here::here("Data/changi_stats.xlsx")) %>% 
  mutate(date = date(local_time))


#summary of daily sum of, and total departures and arrivals per day 

tfc_flow<-changi_stats %>% 
  group_by(date) %>% 
  summarise(arrival_tfc = sum(arrival_tfc),
            departure_tfc = sum(departure_tfc)) %>% 
  mutate(daily_tfc = departure_tfc+ arrival_tfc) %>% 
  mutate(mean_hourly_tfc = daily_tfc/24)
daily_stats_weather<-left_join(tfc_flow, changi_daily_weather)

```



```{r include=FALSE}
changi_hourly_weather<-read_excel(here::here("Data/weather_june18_feb19.xlsx")) 

changi_hourly_weather <-changi_hourly_weather %>% 
  mutate(date = date(local_time))

combined_data<- left_join(changi_hourly_weather, changi_stats) %>% 
  mutate(date=date(local_time))

rainy_days <- daily_stats_weather %>% 
  filter(daily_rainfall_mm>0) 

rainy_date<-rainy_days %>% 
  select(date)
```



```{r include=FALSE}

#combining daily traffic data and weather summaries
daily_stats_weather<-left_join(tfc_flow, changi_daily_weather)


#To check if our hypothesis that weather affects air traffic flows is true, we look at days when there is some rain in Changi
rainy_days <- daily_stats_weather %>% 
  filter(daily_rainfall_mm>0) # filtering out dates when changi experiences some rainfall
#Out of 243 days, 113 days had some rain in Changi - that's almost half

rainy_date<-rainy_days %>% 
  select(date) #filtering out the date component to join after performing MDS.
```


Setting up the distance matrix of weather & air traffic flow during rainy days
```{r}
weather_matrix <- rainy_days %>% 
  select(daily_rainfall_mm, temp_mean_daily, daily_tfc, mean_hourly_tfc) %>%  #selection of 3 dimensions initially to check for sanity
  as.matrix()
```


```{r}
weather_distance<- distances::distances(weather_matrix, normalize = "studentize") %>% 
  as.matrix()

mds_weather <- cmdscale(weather_distance) %>% 
  as_tibble()

mds_weather <- rainy_date %>% 
  select(date) %>% 
  bind_cols(., mds_weather)
```

2.Results from MDS of daily rainfall, traffic flow and daily mean temperature on rainy days in Changi Airport
```{r}
ggplot(mds_weather, aes(x = V1, y = V2)) + geom_text(aes(label= date))
```

##Analysis
However, preliminary results indicate that the distances between the variables dont vary too much and clusters around the (0,0) axis. The outliers are of interest and subsequent data analyses will seek to explore why there is a significant distance between them and the main data clustering.
