# Cluster Analysis


```{r include=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(ggplot2)
library(factoextra)
library(broom)
library(psych)
```


```{r}
#Weather data - July 2018 to February 2019 (daily summaries)
changi_daily_weather <- read_excel(here::here("Data/daily_weather_changi.xlsx")) %>% 
  mutate(date=date(date))

# Changi traffic data - July 2018 to February 2019 (hourly data)
changi_stats <- read_excel(here::here("Data/changi_stats.xlsx"))

#Changi Weather and Traffic data - July 2018 to February 2019 (hourly data)
changi_hourly_weather<-read_excel(here::here("Data/weather_june18_feb19.xlsx")) 


changi_hourly_weather <-changi_hourly_weather %>% 
  mutate(total_tfc = departure_tfc+arrival_tfc)
```



```{r}

hourly_stats_weather <-changi_hourly_weather %>% 
  mutate(hour = hour(local_time)) %>% 
  mutate(date = date(local_time))

```


```{r}

changi_stats <- read_excel(here::here("Data/changi_stats.xlsx")) %>% 
  mutate(daily_tfc = departure_tfc+arrival_tfc) %>% 
  mutate(date = date(local_time))


daily_tfc_flow<-changi_stats %>% 
  group_by(date) %>% 
  summarise(daily_tfc = sum(daily_tfc))

daily_stats_weather <- left_join(changi_daily_weather, daily_tfc_flow)
```


```{r include=F}

rain_hours <-hourly_stats_weather %>% 
  filter(weather_main == "Rain"| weather_main == "Thunderstorm")


#daily_stats_weather <-daily_stats_weather %>% 
#  mutate(weather = if_else(daily_rainfall_mm == 0.0, "NO RAIN", "RAIN"))


```



```{r include = FALSE}
hourly_stats_weather<-hourly_stats_weather %>% 
  mutate(temp = temp/10) %>% 
  mutate(temp_min = temp_min/10) %>% 
  mutate(temp_max = temp_max/10)

daily_stats_weather<- daily_stats_weather %>%
  mutate(daily_wind_speed = wind_speed) %>% 
  mutate(daily_max_wind_speed = max_wind_speed) %>%
  select(-wind_speed, - max_wind_speed)
```




```{r}
data<- left_join(hourly_stats_weather, daily_stats_weather, by = "date")
```



```{r include = FALSE}
hourly_cluster_data<-hourly_stats_weather %>% 
  select(-date, -hour, -weather_id, -weather_description, -weather_main) %>% 
  sample_n(100)

hourly_cluster_data<-hourly_cluster_data %>% 
  column_to_rownames(var = "local_time")
```



```{r include = FALSE}
fa_daily_stats <- daily_stats_weather %>% 
  sample_n(100) %>% 
  column_to_rownames(var = "date") %>%
  principal(nfactors = 4, rotate = "varimax")

fa_daily_stats
```



```{r include = FALSE}
fa_daily_stats[['scores']] %>% 
  unclass() %>% 
  as_tibble(rownames = "date") %>%
  ggplot(aes(x = RC1, y = RC2)) + geom_text(aes(label = date))
```



```{r include=FALSE}
fa_hourly_stats <- hourly_stats_weather %>% 
  sample_n(100) %>% 
  column_to_rownames(var = "local_time") %>% 
  select(-weather_id, -weather_main, -weather_description, -clouds_all, -date) %>% 
  principal(nfactors = 4, rotate = "varimax")

fa_hourly_stats
```
##K-Means Clustering

A k-means clustering is a user-defined clustering algorithm that iteratively changes its centroid based on the lowest total sum of square distances between the centroids and each data point. We use this to help evaluate our visual evaluation of clusters in the factor analysis conducted earlier.  

###Daily Traffic & Weather Data {-}
```{r}
fa_daily_cluster<-fa_daily_stats[['scores']] %>% 
  unclass() %>% 
  as_tibble(rownames = "date")


fa_daily_cluster<-fa_daily_cluster %>% 
  column_to_rownames(var = "date")
```

####Optimal Number of Clusters{-}

```{r}
fviz_nbclust(fa_daily_cluster, kmeans, method = "wss")
```


```{r}
fviz_nbclust(fa_daily_cluster, kmeans, method = "silhouette")
```


```{r}
daily_kmeans_clusters <- kmeans(fa_daily_cluster, centers = 7, nstart = 50)

fviz_cluster(daily_kmeans_clusters, data = fa_daily_cluster)
```

###Hourly Traffic & Weather Data {-}
```{r}
fa_hourly_cluster<-fa_hourly_stats[['scores']] %>% 
  unclass() %>% 
  as_tibble(rownames = "local_time")

fa_hourly_cluster<-fa_hourly_cluster %>% 
  column_to_rownames(var = "local_time")

```

####Optimal Number of Clusters{-}
```{r}
fviz_nbclust(fa_hourly_cluster, kmeans, method = "wss")
```


```{r}
fviz_nbclust(fa_hourly_cluster, kmeans, method = "silhouette")
```


```{r}
hourly_kmeans_clusters <- kmeans(fa_hourly_cluster, centers = 4, nstart = 50)

fviz_cluster(hourly_kmeans_clusters, data = fa_hourly_cluster)
```

##Data Analysis 


